---
title: "A report for `r params$species`"
author:
  - Nithin M^[Indian Institute of Technology-Kharagpur, write2nithinm@iitkgp.ac.in]
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
  pdf_document:
    number_sections: true
params:
  species: Adelie
  category: sex
  
fontsize: 12pt
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r library, message=TRUE, include=FALSE, paged.print=TRUE}
library(gapminder)
library(tidyverse)
library(kableExtra)
library(palmerpenguins)
library(epoxy)
library(glue)
use_epoxy_glue_engine()
data<-penguins
head(data,15)
```
# Data
Let's have a look at the penguins dataset. The dataset contains the details of **`r nrow(data)`** penguins. The dataset contains details for `r n_distinct(data$species)` species of penguins. Let us look at the data for **`r params$species`**.

```{r chunk 1, echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}
library(skimr)
library(stargazer)
library(gt)
library(kableExtra)
library(flextable)


data_species<- data%>% filter(species==params$species) 
data_species%>% head(30) %>% gt()
```

# Summary Statistics
Now that we have the data, Lets see some summary statistics.

```{r summary, echo=FALSE, message=FALSE, warning=FALSE}
library(psych)

st1<-data_species %>% select_if(is.numeric) %>% select(-year) %>% skim() %>% 
        select(Variable=skim_variable, Mean=numeric.mean,
               "Std.Deviation"=numeric.sd,Median=numeric.p50)
st1 %>% gt()
st1<-st1 %>% mutate(Units=c("mm","mm","mm","g"))
```


We observe that
```{glue , sum-species, eval=FALSE}
-  {Variable} has a mean of {round(Mean,3)} {Units}, standard deviation of {round(Std.Deviation,3)} and Median of {Median}.
```

```{glue ref.label="sum-species", data = st1}
```

# Plots
Plots are fundamental in exploratory data analysis and also a visual aid in communication.Let us create some plots

## Box plot
A box and whisker plot—also called a box plot—displays the five-number summary of a set of data. The five-number summary is the minimum, first quartile, median, third quartile, and maximum. In a box plot, we draw a box from the first quartile to the third quartile.

```{r boxplot, echo=FALSE}
library(ggtext)
 bp<-data_species %>% drop_na() %>% 
        ggplot(aes(x=get(params$category),y=flipper_length_mm,fill=get(params$category)))+
        geom_boxplot()+
        coord_flip()+
        labs(title = glue("{params$category} wise variation in flipper length of {params$species}"),x=glue("{params$category}"),
             y="Flipper length (in mm)",
             fill=glue("{params$category}"))+
        geom_jitter(size=0.8,
                    color="dodgerblue",alpha=0.5)+
   ggthemes::theme_clean()
bp
```
```{r check, echo=FALSE, message=FALSE, warning=FALSE}

bp1<-ggplot_build(bp)
bp_stats<-bp1$data[[1]]
if(length(bp_stats$group)==1){
  great<-which(levels(bp1$plot$data[,"island",drop=T])==
                 unique(bp1$plot$data[,"island",drop=T]))
}else
  {
great<-which(bp_stats$middle==max(bp_stats$middle))
}
cat_great<-levels(data_species[,glue("{params$category}"),drop=T])

```
```{r echo=FALSE,comment=NA}

if(params$category=="sex"){
 res<-str_glue("We observe that {cat_great[great]} penguins of {params$species} have greater flipper length")
} else if (params$category=="island"){
res<-str_glue("We observe that  penguins of {params$species} living in {cat_great[great]} island have greater flipper length")
}

```

`r res`

### Density plots
Let us create density plot as well

```{r density,echo=FALSE,message=F,warning=F}
data_species%>% 
        ggplot(aes(x=body_mass_g,
                   fill=get(params$category)))+
        geom_density(
                     size=0.8,
                     alpha=0.4)+
        labs(title=glue("Body mass distribution of {params$species}"),
             subtitle = "Be careful about NA",
             x="Body mass in grams",
             y="Density",
             fill=glue("{params$category}"))+
        theme_bw()+
        theme(title = element_text(face = "italic",
                                   colour="dodgerblue"))+
        scale_y_continuous(labels = function(x) format(x, scientific = FALSE))
```

## Scatter plot

Let us investigate some relationships between the variables.

### Bill length v/s Flipper length by sex
```{r scatter plot, echo=FALSE, message=FALSE, warning=FALSE}
data_species %>% drop_na()%>% ggplot(aes(x=bill_length_mm,
                   y=flipper_length_mm,
                   color=sex))+
        geom_point()+theme_bw()+
        labs(title = glue("Scatter plot for {params$species}"))

```
Let us see some more relationships.

```{r ggpairs, echo=FALSE, message=FALSE, warning=FALSE}
library(GGally)
data_pairs<-data_species%>% drop_na()%>% select(-c(sex,species,island)) %>% 
        ggpairs()+
        labs(title = glue("Different relationships for {params$species}"))+
        theme_bw()
```
### Correlation 
Correlation is a statistical measure that expresses the extent to which two variables are linearly related (meaning they change together at a constant rate).

```{r cor_matrix, echo=FALSE, message=FALSE, warning=FALSE}
library(corrr)
cor_mat<-data_species %>% drop_na() %>% select(-c(sex,species,island)) %>% 
        correlate %>% mutate_if(is.numeric,round,3) 
cor_mat %>% gt()
```
we can observe that the correlation between bill length and flipper length of `r params$species` is `r cor_mat$flipper_length_mm[1]`

Again, picture is more appealing
```{r corr_plot, echo=FALSE, message=FALSE, warning=FALSE}
library(patchwork)
species_plot<-data_species %>% drop_na()%>% select(-c(sex,species,island)) %>% 
        ggcorr(label = T,angle=-45,
               label_alpha = 0.2,
               low="violetred2",
               mid="linen",
               high = "steelblue1")
```
```{r corr_plot_sp, echo=FALSE, message=FALSE, warning=FALSE}
wrap_plots(species_plot)
```

But this is only for `r params$species`. What if we need a plot for all species?
```{r corrPlt full, echo=FALSE, message=FALSE, warning=FALSE}
library(ggpubr)

data_split<-split(data,data$species)
names_species<-names(data_split)
plot3<- map2(data_split,names(data_split),~.x %>%select(-c(sex,species,island)) %>% 
             ggcorr(label = T,angle=-45,
                    label_alpha = 0.2,
                    low="violetred2",
                    mid="linen",
                    high = "steelblue1")+
              labs(title = .y))

```

```{r plots_final, echo=FALSE, message=FALSE, warning=FALSE}
wrap_plots(plot3,nrow = 2)
```

